---
title: "sbdevelopment"
author: "Ben Best"
date: "February 2, 2016"
output:
  html_document:
    toc: true
---

See [raw routing.Rmd](https://raw.githubusercontent.com/bbest/sbdev/gh-pages/routing.Rmd).

## Comparison Test

Survey data:

- location by parcel id
    - AUD: in (n=30)/out of zone (n=120 total)
    - topography: high vs low slope areas determining bikeability
- miles commute to work (by zip code)

Dealing with topography:

- `raster::terrain(dem, opt='slope', unit='radians', neighbors=8)`
- [Santa Barbara, California Coastal Digital Elevation Model](http://www.ngdc.noaa.gov/docucomp/page?xml=NOAA/NESDIS/NGDC/MGG/DEM/iso/xml/603.xml&view=getDataView&header=none) @ 10 - 90m resolution
- [Data Catalog - SB County](http://cosb.countyofsb.org/gis/default.aspx?id=2802)



Test: in/out of AUD ~ commute less/more. control for topography. Goleta/Carp diff't from dtown SB.

Issues:
- `?sample`

## `walkscoreAPI`

Ditched this method because package out of date, and API for transit score not available. Note `eval=F` in Rmd of R chunk.

Links:

* [101 Juana Maria Street, Santa Barbara CA - Walk Score](https://www.walkscore.com/score/101-juana-maria-st-santa-barbara-ca-93103)
* [1319 Chapala Street, Santa Barbara CA - Walk Score](https://www.walkscore.com/score/1319-chapala-st-santa-barbara-ca-93101)
* [Walk Score Methodology](https://www.walkscore.com/methodology.shtml)
* [Bike Score Methodology](https://www.walkscore.com/bike-score-methodology.shtml)
* [Transit Score Methodology](https://www.walkscore.com/transit-score-methodology.shtml)

Let's use the [Walk Score API for web and mobile developers](https://www.walkscore.com/professional/api.php), specifically the [CRAN - Package walkscoreAPI](https://cran.r-project.org/web/packages/walkscoreAPI/).

- Your WS API quota is currently 5,000 calls per day.

```{r walkscoreAPI, eval=F}
library(walkscoreAPI) # install.packages('walkscoreAPI') # 

# added research=yes to URL for getTS to work
getTS = function (x, y, city, state, key, research='yes'){
    city <- gsub(" ", "+", city)
    URL <- paste("http://transit.walkscore.com/transit/score/?lat=", 
        y, "&lon=", x, "&city=", city, "&state=", state, "&wsapikey=", 
        key, "&research=", research, sep = "")

    X <- character(0)
    X <- c(X, scan(file = URL, what = "", sep = "\n", quiet = TRUE))
    string <- X[grep("transit_score", X)]
    string2 <- X[grep("ws_link", X)]
    string3 <- X[grep("description", X)]
    string4 <- X[grep("summary", X)]
    if (length(X) > 0) {
        tscore <- strsplit(string, ": ")
        tscore <- gsub(",", "", tscore[[1]][2])
        tscore <- as.numeric(tscore)
        link <- strsplit(string2, ": ")
        link <- gsub("\"", "", link[[1]][2], fixed = TRUE)
        link <- gsub(", ", "", link)
        desc <- strsplit(string3, ": ")
        desc <- gsub("\"", "", desc[[1]][2], fixed = TRUE)
        desc <- gsub(", ", "", desc)
        sum <- strsplit(string4, ": ")
        sum <- gsub("\"", "", sum[[1]][2], fixed = TRUE)
        sum <- gsub(", ", "", sum)
    }
    else {
        print("error, please check supported cities list")
    }
    object <- list()
    class(object) <- "TransitScore"
    object$transitscore <- tscore
    object$url <- link
    object$description <- desc
    object$summary <- sum
    return(object)
}


ws_key = 'e5bc0a4a03023ef0989dc3b3d27cde6d'
# for (i in 1:nrow(df)){ 

  street = '1319 Chapala St' # df[i,'street']
  city = 'Santa Barbara'     # df[i,'city']
  state = 'CA'
  
  address = sprintf('%s, %s %s', street, city, state)
  xy = ggmap::geocode(address)
  
  # transit score
  ts = getTS(xy$lon, xy$lat, tolower(city), tolower(state), ws_key)
  
  # walk score
  ws = getWS(xy$lon, xy$lat, ws_key)
  

# } # end loop
```

## Google Maps Directions API

The following constructs a URL based on the [Google Maps Directions API](https://developers.google.com/maps/documentation/directions/intro#TravelModes) and processes the JSON return object.

You'll need to [get a Google Maps API Key](https://developers.google.com/maps/documentation/distance-matrix/get-api-key#key) or just use mine.

Note the `*.csv` input and output to the [`data`](https://github.com/bbest/sbdev/tree/gh-pages/data) folder.

```{r gmapsdistance}
suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
  library(stringr)
  library(jsonlite)
})

# set variables: working directory, csv in/out, google maps key
csv_in  = 'data/addresses.csv'
csv_out = 'data/addresses_gmdist.csv'
gm_key = 'AIzaSyCSkbz0pwqF2kPgZ3rReyOlGpSFrqaVhLQ'

# google maps directions api options
departure_time = '2016-02-03 8:00:00' %>% # future. integer in seconds since midnight, January 1, 1970 UTC
  strptime('%Y-%m-%d %H:%M:%S') %>% 
  difftime(strptime('1970-01-01 0:00:00', '%Y-%m-%d %H:%M:%S'), units='secs') %>%
  as.integer()
modes = c('driving','walking','bicycling','transit')
  
# set wd, read addresses, add begin addresses and url columns
a = read_csv(csv_in) %>%
  mutate(
    beg_address = sprintf('%s, %s CA', beg_street, beg_city),
    beg_address_url = beg_address %>%
      str_replace_all(c(' '='+',','='')),
    url = paste0(
      'https://maps.googleapis.com/maps/api/distancematrix/json?',
      'origins=', beg_address_url,
      '&destinations=', end_zip,
      '&key=', gm_key,
      '&departure_time=', departure_time))

# iterate through addresses
for (i in 1:nrow(a)){ 
  #cat(sprintf('%03d: %s -> %s\n', i, a$beg_address_url[i], a$end_zip[i]))
  
  # iterate through modes
  for (mode in modes){

    # input
    url_mode = paste0(a$url[i],'&mode=',mode)
    fld_dist_km = sprintf('%s_dist_km', mode)
    fld_time_min = sprintf('%s_time_min', mode)
    
    # fetch result object from Google
    #cat(sprintf('  %s: %s\n', mode, url_mode))
    o = fromJSON(url_mode, simplifyVector=F)
    
    # interpret result object
    if (o$status != 'OK'){
      cat(sprintf('  %s: %s\n', mode, url_mode))
      cat(sprintf('    ERROR with:\n      %s\n%s', url_mode, str(o)))
    } else{
      #if (i==2) browser() # select(a, -url)
      a[i, fld_dist_km] = o$rows[[1]]$elements[[1]]$distance$value / 1000
      a[i, fld_time_min] = o$rows[[1]]$elements[[1]]$duration$value / 60
    }
    #print(select(a, -url))
  }
  
} # end loop

# write out csv
a %>%
  write_csv(csv_out)

# output table
a %>% 
  select(beg_address, end_zip, ends_with('_km'), ends_with('_min')) %>%
  DT::datatable()
```

## `ggmap::route`

Here we'll use the `ggmap` package to get a detailed route and plot it. Note that you should be able to get the above information by summarizing this table, except `departure_time` is not an available argument.

```{r}
suppressPackageStartupMessages({
  library(ggmap)
})

# vars
mode = 'bicycling'
from = a$beg_address[1]
to   = as.character(a$end_zip[1])

# route
route_df = route(from, to, mode, structure = "route")

# plot
qmap(location='Santa Barbara, CA', zoom=11) +
  geom_path(
    aes(x=lon, y=lat),  colour = "red", size = 1.5,
    data=route_df, lineend = "round")

# table
DT::datatable(route_df)
```
